---
title: transportdata
output:
  github_document:
    number_sections: true
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
[![CRAN status](https://www.r-pkg.org/badges/version/transportdata)](https://CRAN.R-project.org/package=transportdata)
[![R-CMD-check](https://github.com/ITSLeeds/transportdata/workflows/R-CMD-check/badge.svg)](https://github.com/ITSLeeds/transportdata/actions)
<!-- [![Codecov test coverage](https://codecov.io/gh/ITSLeeds/transportdata/branch/master/graph/badge.svg)](https://codecov.io/gh/ITSLeeds/transportdata?branch=master) -->
<!-- badges: end -->


`transportdata` is a package containing transport datasets for teaching, methodological research and software development (e.g. for benchmarking).
Simply put, the aim is for it to be like `spData`, a package used for teaching and demonstrating spatial data methods, but for transport/mobility researchers using R.

# Installation

<!-- You can install the released version of transportdata from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("transportdata") -->
<!-- ``` -->

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ITSLeeds/transportdata")
```  

# Datasets

The datasets included in the package, and code chunks used to create them, are outlined below.
To reproduce the following code, you will need to to have installed a number of packages.
The following packages must be loaded:

```{r, message=FALSE}
library(dplyr)
library(stplanr)
```

## Origin-destination data

OD data is arguably the most fundamental type of data for representing city to global transport systems.
The OD dataset used for this package represents the number of commutes between a subset of desire lines between administrative zones in Leeds, UK.
The dataset was generated as follows:

```{r, eval=FALSE}
od_data = pct::get_od()
# [1] 2402201
nrow(od_data)
od_data_leeds = od_data %>% 
  filter(la_1 == "Leeds") %>% 
  filter(la_2 == "Leeds") 
nrow(od_data_leeds)
# [1] 35463
od_leeds = od_data_leeds %>% 
  filter(all >= 50)
usethis::use_data(od_leeds, overwrite = TRUE)
```


```{r}
dim(transportdata::od_leeds)
```

## Centroid data

```{r}
centroids = pct::get_centroids_ew()
centroids_leeds = centroids %>% 
  filter(stringr::str_detect(string = msoa11nm, pattern = "Leeds")) %>% 
  select(geo_code = msoa11cd) %>% 
  sf::st_transform(4326)
usethis::use_data(centroids_leeds, overwrite = TRUE)
```

```{r}
dim(centroids_leeds)
```


## Route data

```{r}
desire_lines_leeds = od::od_to_sf(od_leeds, centroids_leeds)
routes_leeds_walk = route(l = desire_lines_leeds, route_fun = route_osrm, osrm.profile = "foot")
mapview::mapview(routes_leeds_walk)
saveRDS(routes_leeds_walk, "routes_leeds_walk.Rds")
file.size("routes_leeds_walk.Rds") / 1e6 # 1.6 mb!
rnet_leeds_walk = overline(routes_leeds_walk, "foot")
plot(rnet_leeds_walk["foot"], lwd = rnet_leeds_walk$foot / 500)
piggyback::pb_upload("routes_leeds_walk.Rds")

routes_leeds_bike = route(l = desire_lines_leeds, route_fun = route_osrm, osrm.profile = "bike")
saveRDS(routes_leeds_bike, "routes_leeds_bike.Rds")
file.size("routes_leeds_bike.Rds") / 1e6 # 1.6 mb!
rnet_leeds_bike = overline(routes_leeds_bike, "bicycle")
plot(rnet_leeds_bike["bicycle"], lwd = rnet_leeds_bike$bicycle / 500)
piggyback::pb_upload("routes_leeds_bike.Rds")

routes_leeds_bike = route(l = desire_lines_leeds, route_fun = route_osrm, osrm.profile = "car")
saveRDS(routes_leeds_car, "routes_leeds_car.Rds")
file.size("routes_leeds_car.Rds") / 1e6 # 1.6 mb!
rnet_leeds_car = overline(routes_leeds_car, "car_driver")
plot(rnet_leeds_car["car_driver"], lwd = rnet_leeds_car$car_driver / 500)
piggyback::pb_upload("routes_leeds_car.Rds")

```



# Code of Conduct
  
Please note that the transportdata project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
