---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# transportData

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
[![CRAN status](https://www.r-pkg.org/badges/version/transportData)](https://CRAN.R-project.org/package=transportData)
[![R-CMD-check](https://github.com/ITSLeeds/transportData/workflows/R-CMD-check/badge.svg)](https://github.com/ITSLeeds/transportData/actions)
[![Codecov test coverage](https://codecov.io/gh/ITSLeeds/transportData/branch/master/graph/badge.svg)](https://codecov.io/gh/ITSLeeds/transportData?branch=master)
<!-- badges: end -->

The goal of transportData is to provide example transport datasets for teaching, research and software development.

## Installation

<!-- You can install the released version of transportData from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("transportData") -->
<!-- ``` -->

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ITSLeeds/transportData")
```  

## Datasets

The datasets included to date are outlined below:

```{r}
dim(transportData::od_leeds_50)
```

## Code of Conduct
  
Please note that the transportData project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
