---
title: transportdata
output:
  github_document:
    number_sections: true
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
<!-- [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental) -->
[![CRAN status](https://www.r-pkg.org/badges/version/transportdata)](https://CRAN.R-project.org/package=transportdata)
[![R-CMD-check](https://github.com/ITSLeeds/transportdata/workflows/R-CMD-check/badge.svg)](https://github.com/ITSLeeds/transportdata/actions)
<!-- [![Codecov test coverage](https://codecov.io/gh/ITSLeeds/transportdata/branch/master/graph/badge.svg)](https://codecov.io/gh/ITSLeeds/transportdata?branch=master) -->
<!-- badges: end -->


`transportdata` is a package containing transport datasets for teaching, methodological research and software development (e.g. for benchmarking).
The aim is for it to be for people working with transport data what `spData`, a package used for teaching and demonstrating spatial data methods, is for people working with spatial data.
The package will be especially useful for transport/mobility researchers using R.

# Installation

<!-- You can install the released version of transportdata from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("transportdata") -->
<!-- ``` -->

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ITSLeeds/transportdata")
```  

# Datasets

The datasets included in the package, and code chunks used to create them, are outlined below.
After you have installed the package, the datasets should be available.

```{r}
library(transportdata)
```


To reproduce the following code, you will need to to have installed a number of packages.
The following packages must also be loaded:

```{r, message=FALSE}
library(sf)
library(dplyr)
library(stplanr)
```

## Origin-destination data

OD data is arguably the most fundamental type of data for representing city to global transport systems.
The OD dataset used for this package represents the number of commutes between a subset of desire lines between administrative zones in Leeds, UK.
The dataset was generated as follows:

```{r, eval=FALSE}
od_data = pct::get_od()
# [1] 2402201
nrow(od_data)
od_data_leeds = od_data %>% 
  filter(la_1 == "Leeds") %>% 
  filter(la_2 == "Leeds") 
nrow(od_data_leeds)
# [1] 35463
od_leeds = od_data_leeds %>% 
  filter(all >= 50)
usethis::use_data(od_leeds, overwrite = TRUE)
```


```{r}
dim(transportdata::od_leeds)
```

## Zone data

```{r, eval=FALSE}
z = pct::get_pct_zones("west-yorkshire", geography = "msoa")
zones_leeds = z %>% 
  filter(lad_name == "Leeds") %>% 
  select(geo_code:taxi_other, dutch_slc)
usethis::use_data(zones_leeds)
```


```{r zones}
dim(zones_leeds)
names(zones_leeds)
plot(zones_leeds[5:14])
```


## Centroid data

```{r, eval=FALSE, echo=FALSE}
centroids = pct::get_centroids_ew()
centroids_leeds = centroids %>% 
  filter(stringr::str_detect(string = msoa11nm, pattern = "Leeds")) %>% 
  select(geo_code = msoa11cd) %>% 
  sf::st_transform(4326)
usethis::use_data(centroids_leeds, overwrite = TRUE)
```

```{r centroids}
dim(centroids_leeds)
names(centroids_leeds)
plot(centroids_leeds)
```

## Route data

The route data was generated as follows:

```{r desire}
od_interzonal = od_leeds %>% 
  filter(geo_code1 != geo_code2)
desire_lines_leeds = od::od_to_sf(od_interzonal, centroids_leeds)
desire_lines_leeds_10 = desire_lines_leeds %>% 
  top_n(n = 10, wt = all)
plot(desire_lines_leeds_10)
```

```{r, eval=FALSE}
# get route data for top 10 routes for different modes
routes_leeds_foot_10 = route(l = desire_lines_leeds_10, route_fun = route_osrm, osrm.profile = "foot")
routes_leeds_bike_10 = route(l = desire_lines_leeds_10, route_fun = route_osrm, osrm.profile = "bike")
routes_leeds_car_10 = route(l = desire_lines_leeds_10, route_fun = route_osrm, osrm.profile = "car")
```

```{r, eval=FALSE, echo=FALSE}
usethis::use_data(routes_leeds_foot_10)
usethis::use_data(routes_leeds_bike_10)
usethis::use_data(routes_leeds_car_10)
```

```{r routes-10}
plot(routes_leeds_foot_10$geometry, lwd = 6, col = "green")
plot(routes_leeds_bike_10$geometry, lwd = 4, col = "blue", add = TRUE)
plot(routes_leeds_car_10$geometry, lwd = 2, col = "red", add = TRUE)
```



```{r, eval=FALSE, echo=FALSE}
routes_leeds_foot = route(l = desire_lines_leeds, route_fun = route_osrm, osrm.profile = "foot")
saveRDS(routes_leeds_foot, "routes_leeds_foot.Rds")
piggyback::pb_upload("routes_leeds_foot.Rds")

routes_leeds_bike = route(l = desire_lines_leeds, route_fun = route_osrm, osrm.profile = "bike")
saveRDS(routes_leeds_bike, "routes_leeds_bike.Rds")
piggyback::pb_upload("routes_leeds_bike.Rds")

routes_leeds_car = route(l = desire_lines_leeds, route_fun = route_osrm, osrm.profile = "car")
saveRDS(routes_leeds_car, "routes_leeds_car.Rds")
piggyback::pb_upload("routes_leeds_car.Rds")

```



You can get the full routes as follows:

```{r routes-read}
routes_leeds_foot = get_td("routes_leeds_foot")
routes_leeds_bike = get_td("routes_leeds_bike")
routes_leeds_car = get_td("routes_leeds_car")
```

```{r routes-rnet, out.width="33%", fig.show='hold'}
rnet_leeds_foot = overline(routes_leeds_foot, "foot")
plot(rnet_leeds_foot["foot"], lwd = rnet_leeds_foot$foot / 500)
rnet_leeds_bike = overline(routes_leeds_bike, "bicycle")
plot(rnet_leeds_bike["bicycle"], lwd = rnet_leeds_bike$bicycle / 500)
rnet_leeds_car = overline(routes_leeds_car, "car_driver")
plot(rnet_leeds_car["car_driver"], lwd = rnet_leeds_car$car_driver / 500)
```

## Air quality data

Data on air quality was obtained as follows:

```{r, eval=FALSE}
install.packages("openair")
library(openair)
all_sites = importMeta()
# View(all_sites)
all_sites_sf = sf::st_as_sf(all_sites, coords = c("longitude", "latitude"), crs = 4326)
leeds_sites = all_sites_sf[zones_leeds, ]
headingley = importAURN(site = "LED6", year = 2017:2019, pollutant = c("nox", "no2"))
summary(headingley)
table(headingley$site)
air_quality_leeds_raw = importAURN(site = leeds_sites$code, year = 2017:2019, pollutant = c("nox", "no2"))
air_quality_leeds = right_join(leeds_sites, air_quality_leeds_raw)
usethis::use_data(air_quality_leeds, overwrite = TRUE)
```


```{r airq, out.width="49%", fig.show='hold'}
library(openair)
dim(air_quality_leeds)
names(air_quality_leeds)
table(air_quality_leeds$site)
air_quality_leeds %>% 
  sf::st_drop_geometry() %>% 
  filter(site == "Leeds Centre") %>% 
  TheilSen(pollutant = "no2", ylab = "NO2 (ug/m3)", deseason = TRUE, main = "Leeds Centre")
air_quality_leeds %>% 
  sf::st_drop_geometry() %>% 
  filter(site == "Leeds Headingley Kerbside") %>% 
  TheilSen(pollutant = "no2", ylab = "NO2 (ug/m3)", deseason = TRUE, main = "Leeds Headingley Kerbside")
```


```{r airqmap}
air_quality_summary = air_quality_leeds %>% 
  group_by(site, year = substring(date, 1, 4)) %>% 
  summarise(
    mean_no2 = mean(no2, na.rm = TRUE),
    mean_nox = mean(nox, na.rm = TRUE)
    )
library(tmap)
tm_shape(air_quality_summary, bbox = tmaptools::bb(air_quality_leeds, 1.5)) +
  tm_dots("mean_nox", size = 5) +
  tm_facets(by = "year", nrow = 1) +
  tm_shape(routes_leeds_foot_10) +
  tm_lines() 
```



# Code of Conduct
  
Please note that the transportdata project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
